---
title: "EpiDISH - Epigenetic Dissection of Intra-Sample-Heterogeneity - R package"
author: "Andrew E. Teschendorff, Shijie C. Zheng"
date: "`r Sys.Date()`"
package: "`r pkg_ver('EpiDISH')`"
output:
  BiocStyle::html_document
bibliography: EpiDISH.bib
vignette: >
  %\VignetteIndexEntry{Epigenetic Dissection of Intra-Sample-Heterogeneity - R package}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

# Introduction

The **EpiDISH** package provides tools to infer the proportions of a priori known cell subtypes present in a sample representing a mixture of such cell-types. Inference proceeds via one of 3 methods (Robust Partial Correlations-RPC[@EpiDISH], Cibersort (CBS)[@CBS], Constrained Projection (CP)[@CP]), as determined by user.

For now, the package only includes one whole blood reference of 333 tsDHS-DMCs and 8 blood cell subtypes(B-cells, CD4+ T-cells, CD8+ T-cells, NK-cells, Monocytes, Neutrophils, Eosinophils, and Granulocytes. Note that Granulocytes consist of Neutrophils and Eosinophils.) described in [@EpiDISH]. This referecen dataset was based on 450k DNAm array; however, it could be directly used on both of 450k and EPIC array data. This package is under development and will offer reference-based inference for different tissue types. We will also include more algorithms in the future.


# How to use **EpiDISH** package

Using **EpiDISH** is quite simple. Here we use a small Illumina HumanMethylation450 BeadChip blood dataset(n=2) on GEO as an example.

First, we download the dataset with *getGEO* function in **GEOquery** package and extract beta value matrix.

```{r download, eval=TRUE, echo=T, message=FALSE, warning=FALSE}
require(GEOquery)
require(Biobase)
GSE80559 <- getGEO("GSE80559")
beta.m <- exprs(GSE80559[[1]])
```

Then, we load **EpiDISH** package and the whole blood reference dataset.
```{r load, eval=TRUE, echo=T, message=FALSE, warning=FALSE}
library(EpiDISH)
data(centDHSbloodDMC.m)
```

Notice that *centDHSbloodDMC.m* has 8 columns. Granulocytes consist of Neutrophils and Eosinophils.
So, we only want to inlcude 7 columns(i.e B-cells, CD4+ T-cells, CD8+ T-cells, NK-cells, Monocytes, Neutrophils and Eosinophils) or 6 columns(i.e B-cells, CD4+ T-cells, CD8+ T-cells, NK-cells, Monocytes and Granulocytes). We go ahead and use *epidish* function with *RPC* mode to infer the proportions.
```{r infer, eval=TRUE, echo=T, message=FALSE, warning=FALSE}
ref.m <- centDHSbloodDMC.m[,1:6]
out.l <- epidish(beta.m, ref.m, method = "RPC") 
```

Then, we check the output list. *estF* is the estimated cell fraction matrix. *ref* is the reference centroid matrix used; and *dataREF* is the input data matrix over the probes defined in the reference matrix.
```{r check, eval=TRUE, echo=T, message=FALSE, warning=FALSE}
out.l$estF
dim(out.l$ref)
dim(out.l$dataREF)
```

In this case, all 333 probes in the input reference matrix can be found in the inquiry matrix. So the *ref* is a $333*6$ matrix, while *dataREF* is a $333*2$ matrix. Sometime, in QC step, we might remove bad probes; consequently, not all probes in the reference can be found in inquiry data. By checking *ref* and *dataREF*, we can know the actul probes used to infer the proportions.


# More info about different methods
We compared CP and RPC in [@EpiDISH]. And we also have a review article[@review] which summarized all methods tackling cell heterogeneity for DNAm data. Refers to references section for more details.

# Sessioninfo

```{r sessionInfo, echo=FALSE}
sessionInfo()
```

# References



